---
title: "Mozambique Fatalities"
output: html_notebook
---
# Question: 
Will ACLED record any civilian fatalities in Zimbabwe in October 2017?

# Description: 
The Armed Conflict Location and Event Dataset (ACLED) summarizes press reports on conflict and protest events in Africa and parts of Asia. Outcome will be determined by the "Realtime [Month of Interest] All Africa File" found on the Realtime Data page (ACLED - http://www.acleddata.com/data/realtime-data/). The question will resolve as "yes" if ACLED records any event for which: 'COUNTRY' is the country of interest; 'EVENT_DATE' falls in the period of interest; and 'EVENT_TYPE' is Violence against civilians or Remote violence; and 'FATALITIES' is greater than or equal to 1. Question will be resolved when the relevant data is first released.

# Code
Import libraries and ACLED dataset 
```{r}
library(lubridate) # date manipulations
library(psych)
library(TTR)
library(tm) # text mining
library(SnowballC) # text stemming
library(wordcloud) # wordcloud rendering

df <- read.csv("X:\\Personal Development\\Data Science\\HFC\\zimbabwe_fatalities\\ACLED-Version-7-All-Africa-1997-2016_csv_dyadic-file.csv", header=FALSE, stringsAsFactors = FALSE)

#reconvert first row to header
colnames(df) = df[1, ] # the first row will be the header
df = df[-1, ]          # removing the first row.
```

Subset to relevant Zimbabwe data and convert EVENT_TYPE to character
```{r}
zim_data <- df[which(df$COUNTRY == 'Zimbabwe' & (df$EVENT_TYPE %in% c('Violence against civilians', 'Remote Violence'))), ]
```

Create a date variable consistent with a normal R date 
```{r}
zim_data$DATE <- as.Date(zim_data$EVENT_DATE, format = "%d/%m/%Y")
```


Create a vector of every day from 1/1/2007 to 12/31/2016
```{r}
full_range <- data.frame(DATE = seq(as.Date('2007/1/1'), as.Date('2016/12/31'), 'days'))
```

Merge the two datasets and a flag for dates on which fatalities occurred
```{r}
daily_data <- merge(x = full_range, y = zim_data, by = "DATE", all.x = TRUE)

daily_data$EVENT_FLAG[daily_data$FATALITIES > 0] <- 1

daily_data$EVENT_FLAG[is.na(daily_data$EVENT_FLAG)] <- 0
```

Aggregate the data by day, flagging only dates where a fatality occurred
```{r}
agg_data <- aggregate(daily_data$EVENT_FLAG, by=list(daily_data$DATE), FUN = mean)

agg_data$x[agg_data$x > 0] <- 1
```

Plot the days with events 
```{r}
plot(agg_data$x ~ agg_data$Group.1)
```

## Daily likelihood

Find the daily likelihood of an event
```{r}
daily_likelihood <- mean(agg_data$x)
daily_likelihood
```

So, on any given day there is roughly a 2.5% chance that a protest or riot will occur. 

Find the number of days in October
```{r}
end_date <- as.Date('2017-10-31')

start_date <- as.Date('2017-10-1')

num_days <- as.numeric(end_date) - as.numeric(start_date) + 1 
```

Calculate the likelihood of an event in the month of October
```{r}
compiled_likelihood <- 1 - (1 - daily_likelihood)**num_days
compiled_likelihood
```

## Monthly likelihood

Aggregate the data by month
```{r}
agg_data <- aggregate(daily_data$EVENT_FLAG, by=list(format(daily_data$DATE, "%Y-%m")), FUN = mean)

agg_data$x[agg_data$x != 0] <- 1
```


Calculate monthly likelihood
```{r}
monthly_likelihood <- mean(agg_data$x)
monthly_likelihood
```

Use the average of these likelihoods 
```{r}
(compiled_likelihood + monthly_likelihood)/2
```

## Word Cloud

I want to try to determine common elements among the event descriptions, so I am going to create a wordcloud 

Create a corpus
```{r}
zimCorpus <- Corpus(VectorSource(zim_data$NOTES))
```

Remove punctuation and stopwords from corpus
```{r}
zimCorpus <- tm_map(zimCorpus, removePunctuation)
zimCorpus <- tm_map(zimCorpus, removeWords, stopwords('english'))
```

Perform stemming - words are converted to their stem so as not to duplicate root words that are used in different forms
```{r}
zimCorpus <- tm_map(zimCorpus, stemDocument)
```

Plot the word cloud
```{r}
wordcloud(zimCorpus, max.words = 100, random.order = FALSE, colors=brewer.pal(8, "Dark2"))
```



